---
title: 'SQL'
output: html_document
---

#Filtering Rows and Columns

With SQL, you can filter columns with SELECT and rows with WHERE.  Let's look at an example using THE lahman baseball database.  First, we need to load some packages.
```{r warning=FALSE, message=FALSE}
library(Lahman)
library(sqldf)
```

####Example 1
Suppose we want to see the homerun totals for the 1927 Yankees.  We could write the following:
```{r}
query<-"SELECT playerID,yearID,TEAMid,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```
***

####Example 2
Here, we can see all players from Yankees that have scored at least 40 homeruns.
```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting 
WHERE teamID='NYA' and HR>=40"

sqldf(query)
```
***

####Example 3
Here, we can see those players who had 40 homeruns but didn’t have more than 60 strikeouts.
```{r}
query<-"SELECT playerID,yearID,HR,SO FROM Batting
WHERE HR>40 and SO<60"

sqldf(query)
```
***

####Example 4
We can also see the statistics from 1970’s for Phillies for all players who have had more than 30 homeruns.
```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='PHI' and HR>30 and yearID<=1979 and yearID>=1970"

sqldf(query)
```
***

#Order By
####Example 1
We can also arrange them by descending order of number of homeruns.
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```
***

####Example 2
We can narrow down the search for players that had at least 400 at bats but had less than 10 strikeouts and arrange it by the least amount of strikeouts.
```{r}
query<-"SELECT playerID,yearID,teamID,SO,AB FROM Batting
WHERE SO<10 and AB>=400
ORDER BY SO ASC"

sqldf(query)
```
***


#Aggregation
We can narrow down the output by using aggregation

####Example 1
```{r}
query<-"SELECT playerID, sum(HR) FROM Batting
WHERE playerID='ruthba01'
GROUP BY playerID"

sqldf(query)
```
***

####Example 2
```{r}
query<-"SELECT playerID, sum(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"

sqldf(query)
```
***

#Joining
Joining allows data from multiple tables to be used together by using the commands INNER JOIN and ON.

####Example 1
```{r}
query<-"SELECT nameFirst,nameLast,yearID,teamID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```
***
####Example 2
```{r}
query<-"SELECT nameFirst,nameLast,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
WHERE nameFirst='Bob'
GROUP BY Salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) DESC"

sqldf(query)
```
***
